---
title: "Strava Data"
description: |
  Article on how to effectively scrape and store Strava data using the `targets`
  package
author:
  - name: Julian During
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: cosmo
repository_url: https://github.com/duju211/pin_strava
creative_commons: CC BY
base_url: https://www.datannery.com/posts/strava-data/
editor_options: 
  chunk_output_type: console
---

I am an avid runner and cyclist. For the past couple of years,
I have recorded almost all my activities on some kind of GPS device.

I record my runs with a Garmin device and my bike rides with a Wahoo device,
and I synchronize both accounts on Strava.
I figured that it would be nice to directly access my data from my Strava
account.

In the following text, I will describe the progress to get Strava data into R,
process the data, and then create a visualization of activity routes.

```{r setup, include=FALSE}
source("libraries.R")

Sys.setenv("RSTUDIO_VERSION" = '1.4.1725')
```

You will need the following packages:

```{r, code=read_lines("libraries.R"), eval=FALSE}
```

# Data {.tabset}

The whole data pipeline is implemented with the help of the `targets` package.
You can learn more about the package and its functionalities
[here](https://docs.ropensci.org/targets/).

In order to reproduce the analysis, perform the following steps:

* Clone the repository:
[https://github.com/duju211/pin_strava](https://github.com/duju211/pin_strava)
* Install the packages listed in the `libraries.R` file
* Run the target pipeline by executing `targets::tar_make()` command
* Follow the instructions printed in the console

## Target Plan

We will go through the most important targets in detail.

## OAuth Dance from R

The Strava API requires an ‘OAuth dance’, described below.

### Create an OAuth Strava app

To get access to your Strava data from R, you must first create a Strava API.
The steps are documented on the
[Strava Developer site](https://developers.strava.com/docs/getting-started/).
While creating the app, you’ll have to give it a name. In my case,
I named it `r_api`.

After you have created your personal API, you can find your Client ID and
Client Secret variables in the
[Strava API settings](https://www.strava.com/settings/api).
Save the Client ID as STRAVA_KEY and the Client Secret as STRAVA_SECRET
in your R environment.

<aside>
You can edit your R environment by running `usethis::edit_r_environ()`,
saving the keys, and then restarting R.
</aside>

```
STRAVA_KEY=<Client ID>
STRAVA_SECRET=<Client Secret>
```


The information in `my_sig` can now be used to access Strava data.
Set the `cue_mode` of the target to ‘always’ so that the following API calls
are always executed with an up-to-date authorization token.

## Current authenticated user

Download information about the currently authenticated user.
When preprocessing the data, the columns
`r glue_collapse(tar_read(user_list_cols), sep = ", ", last = " and ")` need
special attention, because they can contain multiple entries and can be
interpreted as list columns.

```{r, code=read_lines("R/active_user.R"), eval=FALSE}
```

In the end there is a data frame with one row for the currently authenticated
user:

```{r, echo=FALSE}
tar_read(df_active_user)
```

## Activities

Load a data frame that gives an overview of all the activities from the data.
Because the total number of activities is unknown, use a while loop.
It will break the execution of the loop if there are no more activities to read.

```{r, code=read_lines("R/read_all_activities.R"), eval=FALSE}
```

The resulting data frame consists of one row per activity.
Make sure that all ID columns have a character format and improve the column
names.

```{r, echo=FALSE}
tar_read(df_act)
```

Extract ids of all activities. Exclude activities which were recorded manually,
because they don't include additional data:

```{r, file="R/rel_act_ids.R"}
```

## Measurements

A ‘stream’ is a nested list (JSON format) with all available measurements of
the corresponding activity.

To get the available variables and turn the result into a data frame,
define a helper function `read_activity_stream`.
This function takes an ID of an activity and an authentication token,
which you created earlier.

Preprocess and unnest the data in this function.
The column `latlng` needs special attention,
because it contains latitude and longitude information.
Separate the two measurements before unnesting all list columns.

```{r, code=read_lines("R/read_activity_stream.R"), eval=FALSE}
```

Do this for every id and save the resulting data frames as `feather` file.
By doing so we can later effectively query the data.

<aside>
The name of the board is determined by the currently logged in user and will
have a different name, if you run the pipeline.
</aside>

# Visualisation

Visualize the final data by displaying the geospatial information in the data.
Join all the activities into one data frame. To do this, get the paths to all
the measurement files:

```{r, file="R/meas_paths.R"}
```

In the final plot every facet is one activity.
Keep the rest of the plot as minimal as possible.

```{r, code=read_lines("R/vis_meas.R"), eval=FALSE}
```

```{r gg_strava, echo=FALSE, fig.height=11, fig.width=11}
tar_read(gg_meas)
```

And there it is:
All your Strava data in a few tidy data frames and a nice-looking plot.
Future updates to the data shouldn’t take too long, because only measurements
from new activities will be downloaded. With all your Strava data up to date,
there are a lot of appealing possibilities for further data analyses of your
fitness data.
