---
title: "Strava Data"
description: |
  Article on how to effectively scrape and store Strava data using the `targets`
  package
author:
  - name: Julian During
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: cosmo
repository_url: https://github.com/duju211/pin_strava
creative_commons: CC BY
base_url: https://www.datannery.com/posts/strava-data/
---

I am an avid runner and cyclist. For the past couple of years,
I have recorded almost all my activities on some kind of GPS device.

I record my runs with a Garmin device and my bike rides with a Wahoo device,
and I synchronize both accounts on Strava.
I figured that it would be nice to directly access my data from my Strava
account.

In the following text, I will describe the progress to get Strava data into R,
process the data, and then create a visualization of activity routes.

```{r setup, include=FALSE}
source("libraries.R")

Sys.setenv("RSTUDIO_VERSION" = '1.4.1725')
df_manifest <- tar_manifest(
  fields = c("name", "command", "pattern", "cue_mode"))
```

You will need the following packages:

```{r, code=read_lines("libraries.R"), eval=FALSE}
```

# Data {.tabset}

The whole data pipeline is implemented with the help of the `targets` package.
You can learn more about the package and its functionalities
[here](https://docs.ropensci.org/targets/).

In order to reproduce the analysis, perform the following steps:

* Clone the repository: [https://github.com/duju211/pin_strava](https://github.com/duju211/pin_strava)
* Install the packages listed in the `libraries.R` file
* Run the target pipeline by executing `targets::tar_make()` command
* Follow the instructions printed in the console

## Target Plan

We will go through the most important targets in detail.

## OAuth Dance from R

The Strava API requires an ‘OAuth dance’, described below.

### Create an OAuth Strava app

To get access to your Strava data from R, you must first create a Strava API.
The steps are documented on the
[Strava Developer site](https://developers.strava.com/docs/getting-started/).
While creating the app, you’ll have to give it a name. In my case,
I named it `r_api`.

After you have created your personal API, you can find your Client ID and
Client Secret variables in the
[Strava API settings](https://www.strava.com/settings/api).
Save the Client ID as STRAVA_KEY and the Client Secret as STRAVA_SECRET
in your R environment.

<aside>
You can edit your R environment by running `usethis::edit_r_environ()`,
saving the keys, and then restarting R.
</aside>

```
STRAVA_KEY=<Client ID>
STRAVA_SECRET=<Client Secret>
```

The function `define_strava_app` shown below creates the OAuth app:

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "my_app", ])
```

```{r, code=read_lines("R/define_strava_app.R"), eval=FALSE}
```

### Define an endpoint

Define an endpoint called `my_endpoint` using the function
`define_strava_endpoint`.

The `authorize` parameter describes the authorization url and the `access`
argument exchanges the authenticated token.

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "my_endpoint", ])
```

```{r, code=read_lines("R/define_strava_endpoint.R"), eval=FALSE}
```

### The final authentication step

Before you can execute the following steps,
you have to authenticate the API in the web browser.

```{r, echo=FALSE}
df_manifest[df_manifest$name == "my_sig", ] %>%
  mutate(cue_mode = "always") %>%
  knitr::kable()
```

```{r, code=read_lines("R/define_strava_sig.R"), eval=FALSE}
```

The information in `my_sig` can now be used to access Strava data.
Set the `cue_mode` of the target to ‘always’ so that the following API calls
are always executed with an up-to-date authorization token.

## Current authenticated user

Get the id of the currently authenticated user:

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "active_user_id", ])
```

Download information about the currently authenticated user.
Map over the `active_user_id` to make sure, that no information is overwritten.
When preprocessing the data, the columns
`r glue_collapse(tar_read(user_list_cols), sep = ", ", last = " and ")` need
special attention, because they can contain multiple entries and can be
interpreted as list columns.

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "df_active_user", ])
```

```{r, code=read_lines("R/active_user.R"), eval=FALSE}
```

In the end there is a data frame with one row for the currently authenticated
user:

```{r, echo=FALSE}
tar_read(df_active_user)
```

## Activities

Load a data frame that gives an overview of all the activities from the data.
Because the total number of activities is unknown, use a while loop.
It will break the execution of the loop if there are no more activities to read.

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "df_act_raw", ])
```

```{r, code=read_lines("R/read_all_activities.R"), eval=FALSE}
```

The resulting data frame consists of one row per activity:

```{r, echo=FALSE}
tar_read(df_act_raw)
```

Make sure that all ID columns have a character format and improve the column
names.

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "df_act", ])
```

```{r, code=read_lines("R/pre_process_act.R"), eval=FALSE}
```

Use `dplyr::pull()` to extract all activity IDs

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "act_ids", ])
```

## Measurements

A ‘stream’ is a nested list (JSON format) with all available measurements of
the corresponding activity.

To get the available variables and turn the result into a data frame,
define a helper function `read_activity_stream`.
This function takes an ID of an activity and an authentication token,
which you created earlier.

The target is defined with dynamic branching which maps over all activity IDs.
Define the `cue_mode` as 'never' to make sure that every target runs exactly
once.

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "df_meas", ])
```

```{r, code=read_lines("R/read_activity_stream.R"), eval=FALSE}
```

Bind the single targets into one data frame:

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "df_meas_all", ])
```

The data now is represented by one row per measurement series:

```{r, echo=FALSE}
tar_read(df_meas_all)
```

Turn the data into a wide format:

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "df_meas_wide", ])
```

```{r, code=read_lines("R/meas_wide.R"), eval=FALSE}
```

In this format, every activity is one row again:

```{r, echo=FALSE}
tar_read(df_meas_wide)
```

Preprocess and unnest the data.
The column `latlng` needs special attention,
because it contains latitude and longitude information.
Separate the two measurements before unnesting all list columns.

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "df_meas_pro", ])
```

```{r, code=read_lines("R/meas_pro.R"), eval=FALSE}
```

After this step, every row is one point in time and every column is a
measurement at this point in time (if there was any activity at that moment).

```{r, echo=FALSE}
tar_read(df_meas_pro)
```

# Visualisation

Visualize the final data by displaying the geospatial information in the data.
Every facet is one activity. Keep the rest of the plot as minimal as possible.

```{r, echo=FALSE}
knitr::kable(df_manifest[df_manifest$name == "gg_meas", ])
```

```{r, code=read_lines("R/vis_meas.R"), eval=FALSE}
```

```{r gg_strava, echo=FALSE}
knitr::include_graphics(tar_read(gg_meas_save))
```

And there it is:
All your Strava data in a few tidy data frames and a nice-looking plot.
Future updates to the data shouldn’t take too long, because only measurements
from new activities will be downloaded. With all your Strava data up to date,
there are a lot of appealing possibilities for further data analyses of your
fitness data.
